---
title: "WordChoice preprocessing"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(cache=TRUE)
library(here)
here::here()
library(plyr)
library(tidyverse)
library(lme4)
library(psych)
library(sciplot)
source(here::here("analysis_scripts","helper","summarizeData.R"))
```


This document recreates the preprocessing steps for the word choice data files.

First, we'll put together all of the data.

### Load and combine data

### Survey Data

```{r}
survey <- read.csv(here::here("raw_data","Clocks_Survey_February_18_2019_processed_wc2.csv"))
survey$subjCode <- as.character(survey$subjCode)
survey$subjCode[survey$subjCode=='109'] <- "p109"
#retain only key demographic columns
survey <- survey %>%
  select(subjCode,Age, Gender, NativeLang, SecondLangYN, SecondLang)
```

### Experiment Data

Load experiment data and combine

```{r}
#Load data experiment 1
d1 <- read.table(here::here("raw_data","Words_v12_05092017.txt"), header=TRUE,sep="\t")
#add experiment version id
d1$version <- "exp1"
#remove angle_1 to angle_8 columns
#These columns are not meaningful bc they mean potentially different things for different participants
#and are potentially confusing
d1 <- subset(d1,select=-c(angle_1,angle_2,angle_3,angle_4,angle_5,angle_6,angle_7,angle_8))
#add missing columns, convert subjCode to character, convert room to character
d1 <- d1 %>% 
  mutate(subjCode = as.character(subjCode),room = as.character(room), labelAssign = NA,windowSize = NA, block.1 = NA, numWheelTurnsUp = NA, numWheelTurnsDown = NA)
#fix mislabeled participant 109
d1$subjCode[d1$subjCode=='109'] <- 'p109'

#Load data experiment 2
d2 <- read.table(here::here("raw_data","Words_v2.2_102417.txt"), header=TRUE,sep="\t")
#add experiment version id
d2$version <- "exp2"
#remove blit to pim columns
#These columns are not meaningful bc they mean potentially different things for different participants
#and are potentially confusing
d2 <- subset(d2,select=-c(blit,clate,gled,gorm,grah, hoon, noobda,pim))
#add missing columns, convert subjCode to character, convert room to character
d2 <- d2 %>% 
  mutate(subjCode = as.character(subjCode),room = as.character(room),windowSize = NA, block.1 = NA, numWheelTurnsUp = NA, numWheelTurnsDown = NA)

#Load data experiment 3
d3 <- read.table(here::here("raw_data","Words_v2.3_022118.txt"), header=TRUE,sep="\t")
#add experiment version id
d3$version <- "exp3"
#remove blit to pim columns
#These columns are not meaningful bc they mean potentially different things for different participants
#and are potentially confusing
d3 <- subset(d3,select=-c(blit,clate,gled,gorm,grah, hoon, noobda,pim))
#add missing columns, convert subjCode to character, convert room to character
d3 <- d3 %>% 
  mutate(subjCode = as.character(subjCode),room = as.character(room),windowSize = NA, block.1 = NA, numWheelTurnsUp = NA, numWheelTurnsDown = NA)

#Load data experiment 4
d4 <- read.table(here::here("raw_data","Words_v2.4_01182020.txt"), header=TRUE,sep="\t",fill=T)
#add experiment version id
d4$version <- "exp4"
#remove blit to pim columns
#These columns are not meaningful bc they mean potentially different things for different participants
#and are potentially confusing
d4 <- subset(d4,select=-c(blit,clate,gled,gorm,grah, hoon, noobda,pim))
#Convert subjCode to character, convert room to character
d4 <- d4 %>% 
  mutate(subjCode = as.character(subjCode),room = as.character(room))
#rename doubled subject id
d4$subjCode <- ifelse(d4$subjCode=="256"&d4$room=="SEIDENBERG_PC4","256participant2", as.character(d4$subjCode))

#combine all data
d <- bind_rows(d1,d2,d3,d4)

```


### Exclude subjects

Three subjects excluded from version 3

```{r}
#exclude subjects
excludeList <- c("323","328","352","357")
d <- subset(d, !(subjCode %in% excludeList))
```

### Add trial counters 

Track trial numbers overall and by block accurately (these are useful variables that were not originally in the manuscript).

```{r}
d$overallTrialNum <- 0
d$nameBlockNum <- 0
d$pairLearnBlockNum <- 0

for (curSubj in unique(d$subjCode)) {
  #print(curSubj)
  nameBlockNum <- 1
  pairLearnBlockNum <- 1
  nameTrialNum <- 0
  pairLearnTrialNum=0
  
  if (d$version[d$subjCode==curSubj][1]=="exp4") {
    name_threshold <-  120
  } else {
    name_threshold <-  20
  }
  
  for (trialNum in 1: length(d$subjCode[d$subjCode==curSubj])) {
    #print(trialNum)
    d$overallTrialNum[d$subjCode==curSubj][trialNum] <-  trialNum
    if (d$trialType[d$subjCode==curSubj][trialNum]=="name"&d$overallTrialNum[d$subjCode==curSubj][trialNum]>name_threshold) {
      nameTrialNum <- nameTrialNum+1
    }
    if (d$trialType[d$subjCode==curSubj][trialNum]=="pairLearn") {
      pairLearnTrialNum <- pairLearnTrialNum+1
    }
    d$nameTrialNum[d$subjCode==curSubj][trialNum] <-  ifelse(d$trialType[d$subjCode==curSubj][trialNum]=="name"&d$overallTrialNum[d$subjCode==curSubj][trialNum]>name_threshold,nameTrialNum,NA)
    d$pairLearnTrialNum[d$subjCode==curSubj][trialNum] <-  ifelse(d$trialType[d$subjCode==curSubj][trialNum]=="pairLearn",pairLearnTrialNum,NA)
    d$nameBlockNum[d$subjCode==curSubj][trialNum] <-  ifelse(d$trialType[d$subjCode==curSubj][trialNum]=="name"&d$overallTrialNum[d$subjCode==curSubj][trialNum]>name_threshold,nameBlockNum,NA)
    d$pairLearnBlockNum[d$subjCode==curSubj][trialNum] <-  ifelse(d$trialType[d$subjCode==curSubj][trialNum]=="pairLearn",pairLearnBlockNum,NA)
    if (d$trialType[d$subjCode==curSubj][trialNum]=="name"&nameTrialNum%%8==0&d$overallTrialNum[d$subjCode==curSubj][trialNum]>name_threshold) {
      nameBlockNum <- nameBlockNum+1
    }
    if (pairLearnTrialNum%%20==0&d$trialType[d$subjCode==curSubj][trialNum]=="pairLearn") {
      pairLearnBlockNum <- pairLearnBlockNum+1
    }
  }
}
```

### Add columns tracking subject-specific dictionary

```{r}
for (curSubj in unique(d$subjCode)) {
  #words
  d$word1[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==1])[1]
  d$word2[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==2])[1]
  d$word3[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==3])[1]
  d$word4[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==4])[1]
  d$word5[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==5])[1]
  d$word6[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==6])[1]
  d$word7[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==7])[1]
  d$word8[d$subjCode==curSubj] <-  as.character(d$target[d$block=="init"&d$subjCode==curSubj&d$trial==8])[1]
  
  #frequency of words
  d$freq1[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==1])[1]
  d$freq2[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==2])[1]
  d$freq3[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==3])[1]
  d$freq4[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==4])[1]
  d$freq5[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==5])[1]
  d$freq6[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==6])[1]
  d$freq7[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==7])[1]
  d$freq8[d$subjCode==curSubj] <-  as.character(d$freq[d$block=="init"&d$subjCode==curSubj&d$trial==8])[1]
  
  #corresponding angles
  d$angle1[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==1][1]
  d$angle2[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==2][1]
  d$angle3[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==3][1]
  d$angle4[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==4][1]
  d$angle5[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==5][1]
  d$angle6[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==6][1]
  d$angle7[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==7][1]
  d$angle8[d$subjCode==curSubj] <-  d$angle[d$block=="init"&d$subjCode==curSubj&d$trial==8][1]
}
```


### Trim whitespace in label

Trim leading and trailing whitespace in typed label responses.

In order to avoid oversights and maintain consistency with past code, this *overwrites* the column "label", while also creating a new label_raw column to make the original typed response recoverable.

```{r}
#save label responses in a new column
d$label_raw <- d$label
#trim trailing and leading whitespace
d$label <- trimws(as.character(d$label))
```

### Add more substantive variables for assessing final test performance

```{r}
#store original isRight column from the experiment
d <- d %>%
  mutate(isRight_old=isRight)
#evaluate correct responses for test blocks (training pair learning and test; final timed and untimed test)
#also correcting the isRight column for the final name check block, after cleaning whitespace (which was not done in the experiment)
d$isRight <- ifelse(d$block=="test_x",ifelse(d$label==as.character(d$nearbyLabel1),1,0),
                 ifelse(d$block=="init",ifelse(d$label==as.character(d$target),1,0),
                        ifelse(d$block=="name_check",ifelse(d$label==as.character(d$target),1,0),d$isRight)))

# frequency info
d$freq <- ifelse(d$block=="test_x",as.character(d$nearbyFreq1),as.character(d$freq))
```

### Add variables for assessing test performance, defining neighboring angles systematically

```{r}
# Assign 'left' angle to leftmost clockwise position (smallest angle; except, 330 > 15) OR to the exact angle
d$leftAngle <- ifelse(is.na(d$nearbyAngle2),d$nearbyAngle1,
                      ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), 330,
                             ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), 330,
                                    ifelse(d$nearbyAngle1<d$nearbyAngle2,d$nearbyAngle1,d$nearbyAngle2))))
d$rightAngle <- ifelse(is.na(d$nearbyAngle2),NA,
                       ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), 15,
                              ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), 15,
                                     ifelse(d$nearbyAngle1<d$nearbyAngle2,d$nearbyAngle2,d$nearbyAngle1))))

# give word label for the defined "left" and "right" angle.
d$leftLabel <- ifelse(is.na(d$nearbyAngle2),as.character(d$nearbyLabel1),
                      ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), as.character(d$nearbyLabel1),
                             ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), as.character(d$nearbyLabel2),
                                    ifelse(d$nearbyAngle1<d$nearbyAngle2,as.character(d$nearbyLabel1),as.character(d$nearbyLabel2)))))
d$rightLabel <- ifelse(is.na(d$nearbyAngle2),NA,
                       ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), as.character(d$nearbyLabel2),
                              ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), as.character(d$nearbyLabel1),
                                ifelse(d$nearbyAngle1<d$nearbyAngle2,as.character(d$nearbyLabel2),
                                       as.character(d$nearbyLabel1)))))

#Given angle values, what are the labels assigned these angles?
#Identify frequency of left and right word labels
d$leftFreq <- ifelse(is.na(d$nearbyAngle2),as.character(d$nearbyFreq1),
                     ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), as.character(d$nearbyFreq1),
                            ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), as.character(d$nearbyFreq2),
                                   ifelse(d$nearbyAngle1<d$nearbyAngle2,as.character(d$nearbyFreq1),
                                          as.character(d$nearbyFreq2)))))
d$rightFreq <- ifelse(is.na(d$nearbyAngle2),NA,
                      ifelse((d$nearbyAngle1==330 &d$nearbyAngle2==15), as.character(d$nearbyFreq2),
                             ifelse((d$nearbyAngle1==15 &d$nearbyAngle2==330), as.character(d$nearbyFreq1),
                                    ifelse(d$nearbyAngle1<d$nearbyAngle2,as.character(d$nearbyFreq2),
                                           as.character(d$nearbyFreq1)))))

# Each trial target is one of any potential combination of LF and HF neighbors.
#add trial condition labels: HF,LF,HF_LF, LF_HF, LF_LF, HF_HF
d$conditionType <- ifelse(is.na(d$nearbyAngle2)&d$leftFreq=="hf","HF",
                          ifelse(is.na(d$nearbyAngle2)&d$leftFreq=="lf","LF",
                                 ifelse(d$leftFreq=="hf"&d$rightFreq=="lf", "HF_LF",
                                        ifelse(d$leftFreq=="lf"&d$rightFreq=="hf", "LF_HF",
                                               ifelse(d$leftFreq=="lf"&d$rightFreq=="lf", "LF_LF","HF_HF")))))

# There are two overall groups that the conditions could fall into.
# group trial condition labels: different, same
d$conditionTypeComb <- ifelse((d$leftFreq=="hf"&d$rightFreq=="lf")|(d$leftFreq=="lf"&d$rightFreq=="hf"), "different","same")

# the hf word can be the left or the right angle.
# identify, out of left and right, what the hf word label is
d$highFreqLabel <- ifelse(d$conditionType=="HF",d$leftLabel,
                          ifelse(d$conditionType=="LF",NA,
                                 ifelse(d$leftFreq=="hf"&d$rightFreq=="lf", d$leftLabel,
                                        ifelse(d$leftFreq=="lf"&d$rightFreq=="hf", d$rightLabel,
                                               ifelse(d$leftFreq=="lf"&d$rightFreq=="lf", NA,d$leftLabel)))))

# Evaluates distance of trial angle to left angle 
d$angleDiff <- ifelse(d$leftAngle==330 & d$angle<=15,abs(d$angle+360-d$leftAngle),abs(d$angle - d$leftAngle))

# participants could give a neighboring or wrong label. What did they provide?
#add variable to track if choice was one of two surrounding angles
d$rightAngleChoice <- ifelse(d$label==d$rightLabel,1,
                             ifelse(d$label==d$leftLabel,0,NA))

#Out of the trials where 1 of the 2 labels were given, did they choose an hf label?
#add variable to track if choice was high frequency angle
d$highFreqChoice <- ifelse(d$label==d$highFreqLabel,1,
                          ifelse(is.na(d$rightAngleChoice),NA,0))

#track difference from high frequency label
d$angleDiffHighFreq <- ifelse(is.na(d$highFreqLabel),NA,
                             ifelse(d$highFreqLabel==d$leftLabel,d$angleDiff,45-d$angleDiff))

#add general variable to track whether label is a word in the dictionary
d$valid_label <- ifelse(d$label %in% c(unique(d$word1),unique(d$word2),unique(d$word3),unique(d$word4),unique(d$word5),unique(d$word6),unique(d$word7),unique(d$word8)),1,0)
```

### Add key parameters for main HF/LF analysis

```{r}
#HF trial
d$hfTrial <- ifelse((d$angleDiff >= 22.5 & d$conditionType=="LF_HF") | (d$angleDiff <= 22.5 & d$conditionType=="HF_LF"),.5,
                 ifelse((d$angleDiff >= 22.5 & d$conditionType=="HF_LF") | (d$angleDiff <= 22.5 & d$conditionType=="LF_HF"),-.5,NA))
#determine type of choice made
d$listChoice <- ifelse(is.na(d$hfTrial),NA,
                      ifelse(as.character(d$label) == as.character(d$leftLabel)|as.character(d$label)== as.character(d$rightLabel),1,0))

#In all test trials, did they choose an hf label?
#add variable to track if choice was high frequency angle
d$highFreqChoiceGeneral <- ifelse(as.character(d$label)==as.character(d$highFreqLabel),1,0)
d$lowFreqLabel <- ifelse(is.na(d$hfTrial),NA,
                         ifelse(as.character(d$highFreqLabel)==as.character(d$rightLabel),as.character(d$leftLabel),
                               as.character(d$rightLabel)))

d$targetLabel <- ifelse(is.na(d$hfTrial),NA,
                       ifelse(d$hfTrial==-0.5,as.character(d$lowFreqLabel), as.character(d$highFreqLabel)))
d$nearestLabel <- ifelse(is.na(d$leftLabel),NA,                ifelse(d$angleDiff<22.5,as.character(d$leftLabel),
                         as.character(d$rightLabel)))

#add information about string length
d <- d %>%
  mutate(
    targetLabel_length=str_length(targetLabel),
    nearestLabel_length=str_length(nearestLabel))


#determine if match choice.
d$matchChoice=ifelse(is.na(d$hfTrial),NA,
                     ifelse(d$listChoice==0,0,
                            ifelse(d$hfTrial==0.5&d$highFreqChoiceGeneral==1,1,
                                   ifelse(d$hfTrial==-0.5&d$highFreqChoiceGeneral==0,1,0))))

d$angleDiffFromMatch <-  ifelse(d$angleDiff<22.5,d$angleDiff,45-d$angleDiff)

#finer choice classification
d$choiceKind <- ifelse(is.na(d$hfTrial),NA,
                    ifelse(d$listChoice==0,"other",
                           ifelse(d$hfTrial==0.5&d$matchChoice==1,"high frequency",
                                  ifelse(d$hfTrial==-0.5&d$matchChoice==0,"high frequency","low frequency"))))
```

### Add additional centered variables

```{r}
#center test block based on low vs. high ambiguity
d$blockC <- ifelse((d$block=="test_a"|d$block=="clear"|d$block=="clear2"),-0.5,
                ifelse(d$block=="test_b"|d$block=="ambig",0.5,NA))

#angle difference centered by subject
d <- d %>%
  group_by(subjCode) %>%
  mutate(angleDiffFromMatchC=angleDiffFromMatch-mean(angleDiffFromMatch,na.rm=T))

```

### Add information about final accuracy for each label on choice trials

```{r}
#create dictionary of accuracy on final trials for each subject
finalAccuracyDict <-  subset(d, trialType=="finalName",select=c(version,subjCode,target,isRight))

#integrate final accuracy with choice trials
d$finalAccuracyNearbyLabel1 <- NA
d$finalAccuracyNearbyLabel2 <- NA
for (i in 1:length(d$subjCode)) {
  if (d$trialType[i]=="test") {
    #print(i)
    d$finalAccuracyNearbyLabel1[i] <-  finalAccuracyDict$isRight[finalAccuracyDict$version==as.character(d$version[i])&finalAccuracyDict$subjCode==as.character(d$subjCode[i])&finalAccuracyDict$target==as.character(d$nearbyLabel1[i])]
    d$finalAccuracyNearbyLabel2[i] = ifelse(is.na(d$nearbyLabel2[i]),NA,finalAccuracyDict$isRight[finalAccuracyDict$version==as.character(d$version[i])&finalAccuracyDict$subjCode==as.character(d$subjCode[i])&finalAccuracyDict$target==as.character(d$nearbyLabel2[i])])
  }
}

d <- d %>%
  mutate(finalAccuracyNearbyLabels=(finalAccuracyNearbyLabel1+finalAccuracyNearbyLabel2)/2)
```

### Combine experiment data with survey data

```{r, warning=F}
#merge with survey data
d <- merge(d,survey)

#fix Age variable where needed
d <- d %>%
  mutate(Age = case_when(
    subjCode=="p111" ~ 19,
    subjCode=="445" ~ 18,
    TRUE ~ as.numeric(as.character(Age))
  ))
```

### Select final columns

```{r}
d <- d %>%
  select(
    version,
    expVersion,
    subjCode,
    Age,
    Gender,
    NativeLang,
    SecondLangYN,
    SecondLang,
    seed,
    angleAssign,
    labelAssign,
    room,
    windowSize,
    date_time,
    dateStr,
    block,
    blockC,
    trialType,
    overallTrialNum,
    trial,
    trialNumber,
    nameBlockNum,
    pairLearnBlockNum,
    nameTrialNum,
    pairLearnTrialNum,
    word1,
    word2,
    word3,
    word4,
    word5,
    word6,
    word7,
    word8,
    freq1,
    freq2,
    freq3,
    freq4,
    freq5,
    freq6,
    freq7,
    freq8,
    angle1,
    angle2,
    angle3,
    angle4,
    angle5,
    angle6,
    angle7,
    angle8,
    freq,
    conditionType,
    conditionTypeComb,
    hfTrial,
    target,
    angle,
    nearestLabel,
    nearestLabel_length,
    targetLabel,
    targetLabel_length,
    highFreqLabel,
    lowFreqLabel,
    left,
    right,
    nearbyLabel1,
    nearbyAngle1,
    nearbyFreq1,
    nearbyLabel2,
    nearbyAngle2,
    nearbyFreq2,
    leftAngle,
    rightAngle,
    leftLabel,
    rightLabel,
    leftFreq,
    rightFreq,
    angleDiff,
    angleDiffFromMatch,
    angleDiffFromMatchC,
    angleDiffHighFreq,
    updateScore,
    feedback,
    response,
    label,
    label_raw,
    isRight_old,
    isRight,
    rt,
    score,
    numWheelTurnsUp,
    numWheelTurnsDown,
    valid_label,
    listChoice,
    rightAngleChoice,
    highFreqChoice,
    highFreqChoiceGeneral,
    matchChoice,
    choiceKind,
    finalAccuracyNearbyLabel1,
    finalAccuracyNearbyLabel2,
    finalAccuracyNearbyLabels
  )
```


### Save the final data version
```{r}
#write to processed data folder
write.csv(d,here::here("processed_data","Words_final_preprocessed.csv"),row.names=F)
write.csv(subset(d,version=="exp1"), here::here("processed_data","Words_v12_final_preprocessed.csv"),row.names=F)
write.csv(subset(d,version=="exp2"), here::here("processed_data","Words_v2_final_preprocessed.csv"),row.names=F)
write.csv(subset(d,version=="exp3"), here::here("processed_data","Words_v3_final_preprocessed.csv"),row.names=F)
write.csv(subset(d,version=="exp4"), here::here("processed_data","Words_v4_final_preprocessed.csv"),row.names=F)
```

### Session Info

```{r}
sessionInfo()
```